---
title: "FA7 Exponential Distribution"
author: "Baybayon, Darlyn Antoinette B."
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 1. Records show that job submissions to a computer center have a Poisson distribution with an average of four per minute.

Let T be the time in minutes between submissions. 

P(T < 0.25) = P (the time between submissions is less than 0.25 minute
(15 seconds)) 

P(T > 0.5) = P (the time between submissions is greater than 0.5 minute
(30 seconds)) 

P(0.25 ≤ T ≤ 1) = P (the time between submissions is between 0.25 and
1 minute)

With lambda = 4 per minute, use R to obtain:

#### (a) P(T ≤ 0.25) = P(time between submissions is at most 15 seconds);

```{r}
lambda <- 4
time_int <- 0.25

prob_lt_025 <- pexp(time_int, lambda) 
prob_lt_025
```

#### (b) P(T > 0.5) = P(time between submissions is greater than 30 seconds);

```{r}
time_int <- 0.5

prob_lt_05 <- 1- pexp(time_int, lambda) 
prob_lt_05

```

#### (c) P(0.25 < T < 1) = P(time between submissions is between 15 seconds and 1 minute).
```{r}
time_int <- 0.25
up_lim <- 1

prob_bw_025_1 <- pexp(up_lim, lambda) - pexp(time_int, lambda) 
prob_bw_025_1
```



### 3.  The average rate of job submissions in a computer center is 2 per minute. 
If it can be assumed that the number of submissions per minute has a Poisson
distribution, calculate the probability that:

#### (a)  more than two jobs will arrive in a minute; P(X>2) = 1 - P(X≤2)

```{r}
lambda <- 2
1-ppois(2, lambda)

```

#### (b)  at least 30 seconds will elapse between any two jobs; P(X >= 0.5)

```{r}
1- pexp(0.5, lambda) 
```

#### (c) less than 30 seconds will elapse between jobs; P(X<0.5)
```{r}
pexp(0.5, lambda) 
```

#### (d) a job will arrive in the next 30 seconds, if no jobs have arrived in the last 30 seconds
P(X<=0.5 | X> 0.5) = P(X<0.5)

```{r}
pexp(0.5, lambda) 
```



### 7.  A website receives an average of 15 visits per hour, which arrive following a Poisson distribution.


#### (a) Calculate the probability that at least 10 minutes will elapse without a visit. P(X>0.167)
```{r}
lambda <- 15
1- pexp(0.167, lambda)

```
#### (b) What is the probability that in any hour, there will be less than eight visits?
```{r}
ppois(7, lambda)
```

#### (c) Suppose that 15 minutes have elapsed since the last visit, what is the probability that a visit will occur in the next 15 minutes.
P(X<0.25)

```{r}
pexp(0.25, lambda)
```

#### (d) Calculate the top quartile, and explain what it means.

```{r}
qpois(0.75, lambda)
```

This means that in 75% of the hours, the website shall have 18 or less visits each hour and only 25% of any hour will the website have more than 18 visits. 
