---
title: "FA6 Geometric Distribution"
author: "Baybayon, Darlyn Antoinette B."
output: pdf_document
  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

### I. Geometric Distribution. Provide an R code for the geometric distribution. The geometric distribution is a probability distribution that models the number of trials required to achieve the first success in a sequence of Bernoulli trials, where each trial has a constant probability of success.

#### 1. Set the probability of success:  p <- 0.2

```{r}
p <- 0.2


```

#### 2. Generate 1000 random variables from the geometric distribution.

```{r}
x <- rgeom(1000, p)
```

#### 3. Calculate some basic statistics: mean_x <- mean(x), var_x <- var(x), sd_x <- sd(x)

```{r}
mean_x <- mean(x)
var_x <- var(x)
sd_x <- sd(x)

skew_x <- moments::skewness(x)
kurt_x <- moments::kurtosis(x)
```

#### 4. Print the results in item 3 with the following output (string): 

```{r, echo=FALSE}
print(paste("Number of trials required to achieve first success: ", round(1/p)))
print(paste("Mean:", round(mean_x, 2)))
print(paste("Variance:", round(var_x, 2)))
print(paste("Standard Deviation:", round(sd_x,2)))
print(paste("Skewness:", round(skew_x,2)))
print(paste("Kurtosis:", round(kurt_x, 2)))

```

#### 5. Plot the histogram of the results.

```{r, fig.align='center'}
par(mfrow = c(1,2))

plot(table(x+1)/1000, 
     xlab = "No. of trials to success", 
     ylab = "Relative Frequency", 
     type = "h")
plot(x+1, dgeom(x, p, log = FALSE), 
     xlab = "No. of trials to success", 
     ylab = "Probability", 
     type = "h")


```



### II.Hypergeometric Distribution.  Consider a plant manufacturing IC chips of which 10% are expected to be defective. The chips are packed in boxes for export. Before transportation, a sample is drawn from each box. Estimate the probability that the sample contains more than 10% defectives, when:

#### 1. A sample of 10 is selected from a box of 40;

```{r}
p <- 0.10
n <- 10
N <- 40
M <- p*N
m <- p*M

prob_N40 <- 1- phyper(1, M, N-M, n)
prob_N40
```
#### 2. A sample of 10 is selected from a box of 5000.

```{r}
N2 <- 5000
M2 <- p*N2
m2 <- p*M2

prob_N5000 <- 1- phyper(1, M2, N2-M2, n)

prob_N5000

```

